---
title: "SDS 192 Midterm Exam"
subtitle: "Professor Lindsay Poirier"
format: pdf
editor: source
knitr:
  opts_chunk: 
    message: false
    warning: false
---

## Name:


```{r}
#| include: false

library(tidyverse)
nyc_squirrel_census <- read.csv("https://data.cityofnewyork.us/resource/ej9h-v6g2.csv")
```
\noindent\textbf{Instructions:}
\begin{itemize}
\item This midterm lasts 75 minutes.
\item This exam is closed-book, to be individually completed and without the aid of the internet or mobile phones. You may reference the `R` cheatsheets, along with any supplementary material I have provided.
\item If there are any areas of the exam that are ambiguous to you, please note that on the exam, state your assumptions, and use your best judgment.
\end{itemize}

Several questions in this exam will reference an example dataset. That dataset is the 2018 Central Park Squirrel Census. I have provided you with the documentation for that dataset. Please note that you will be working specifically with the **Hectare Data**. 

1. What is the unit of observation in the example dataset?

\vspace{5cm}

2. Identify one nominal, one ordinal, one discrete numeric variable in the example dataset.

\vspace{4cm}

3. What are the characteristics of a data frame in `R`? What are the characteristics of a vector? How are they related?

\vspace{4cm}

4. What is the chief difference between a barplot and a histogram?

\vspace{4cm}

5. We saw in class that there are two ways to create a barplot using `ggplot`: either using `geom_bar()` or
using `geom_col()`. Illustrate the need for these two different ways using simple data frame examples.

\vspace{6cm}

6. What is overplotting? What are two ways that we can deal with overplotting?

\vspace{4cm}

7. When would we use each of the following types of color palettes on a plot: a qualitative palette, a sequential palette, and a divergent palette?

\vspace{4cm}

8. Given the example dataset, what would be an appropriate y-axis label for the following plot?

```{r}
nyc_squirrel_census |>
  ggplot(aes(x = number_of_squirrels)) +
  geom_histogram(binwidth = 2) +
  labs(x = "Number of Squirrels",
       y = "________________________")
```

9. Develop a descriptive title and labels for the following plot.

```{r}
library(lubridate)
nyc_squirrel_census |>
  mutate(date = mdy(date)) |>
  ggplot(aes(x = date, fill = hectare_conditions)) +
  geom_bar(position = "fill") +
  labs(title = "___________________________________",
       x = "________________",
       y = "________________",
       fill = "____________________")
```

10. What is the scale of the x-axis on the previous plot? What is the scale of the y-axis?

\vspace{4cm}

11. According to Tufte, what is a plot's lie factor?

\vspace{4cm}

12. Approximately what is the interquartile range represented on the following plot?

```{r}
nyc_squirrel_census |>
  ggplot(aes(x = total_time_of_sighting)) +
  geom_boxplot()
```

\vspace{4cm}

13. Two students were working on the same lab assignment and tried to edit the same line in the file. The merge conflict below appeared. What should Student 2 do to ensure that their version of the code gets pushed to the remote repository?

```
<<<<<<< HEAD:lab1.Rmd
#Student 1's code
ggplot(lobby_chicago, aes(x = ACTION)) + geom_bar()
-------
#Student 2's code
ggplot(lobby_chicago, aes(x = PURPOSE)) + geom_bar()
>>>>>>> hash
```

\vspace{4cm}

14. Fill in the following statement with the correct `git` vocabulary. 

`git __________` downloads a remote repository to your local computer. `git __________` fetches changes to code from a remote repository to your local machine. __________ involves selecting files we wish to commit changes on. `git __________` saves a snapshot of the current version of code to the git version control. `git __________` uploads changes to a remote repository. 

15. What is branching? Outline a collaborative GitHub workflow that engages branching to incorporate multiple changes into a final document. 

\vspace{4cm}

16. In what hectares and on what dates and shifts did anonymous sighter 37 count? Write pseduo-code to produce the following data frame.

```{r}
#| echo: false

nyc_squirrel_census |>
  filter(anonymized_sighter == 37) |>
  select(hectare, shift, date, anonymized_sighter)
```

\vspace{6cm}

17. What would be the difference in outputs between the following two code chunks?

```{r}
#| eval: false

mean(nyc_squirrel_census$total_time_of_sighting, na.rm = TRUE)

nyc_squirrel_census |>
  summarize(mean_time_of_sighting = mean(total_time_of_sighting, na.rm = TRUE))
```

\vspace{6cm}

18. In which 10 counting sessions where the most squirrels sighted per minute? Return for me their hectares, shifts, dates, and sightings per minute. Write pseudo-code to produce the following data frame.

```{r}
#| echo: false

nyc_squirrel_census |>
  mutate(sightings_per_minute = number_of_squirrels/total_time_of_sighting) |>
  arrange(desc(sightings_per_minute)) |>
  select(hectare, shift, date, sightings_per_minute) |>
  head(10)
```

\vspace{6cm}

19. In which 5 hectares were the most total squirrels squirrels sighted? How much time was spent sighting in each of these hectares? Write pseduo-code to produce the following data frame.

```{r}
#| echo: false

nyc_squirrel_census |>
  group_by(hectare) |>
  summarize(total_squirrels = sum(number_of_squirrels, na.rm = TRUE),
            total_time_hectare = sum(total_time_of_sighting, na.rm = TRUE)) |>
  arrange(desc(total_squirrels)) |>
  head(5)
```

\vspace{6cm}

20. What was the total, mean, median time spent sighting squirrels in AM vs. PM counting sessions, and how many times was the total time spenting counting recorded as `NA`? Write pseduo-code to produce the following data frame.


```{r}
#| echo: false

nyc_squirrel_census |>
  group_by(shift) |>
  summarize(total_time = sum(total_time_of_sighting, na.rm = TRUE),
            mean_time = mean(total_time_of_sighting, na.rm = TRUE),
            median_time = median(total_time_of_sighting, na.rm = TRUE), 
            na_time = sum(is.na(total_time_of_sighting)))
```

\vspace{6cm}
